Commutative character sheaves and geometric types for supercuspidal representations

### 2020-10-20 referee_response.tex

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\kbar}{\overline{k}}
\newcommand{\A}{\mathbb{A}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Qp}{\Q_p}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}_{\rm div}}
\newcommand{\Zp}{\Z_p}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Fp}{\F_p}
\newcommand{\Fq}{\F_q}
\newcommand{\Fqbar}{\overline{\F}_q}
\newcommand{\Fqm}{\F_{q^m}}
\newcommand{\Fqn}{\F_{q^n}}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\dom}{\operatorname{dom}}
\newcommand{\Spec}{\operatorname{Spec}}

\title{Response to Referee -- \\ Commutative character sheaves and geometric types for supercuspidal representations}
\author{Clifton Cunningham and David Roe}
\date{\today}

\begin{document}
\maketitle

We would like to thank the referee for a thorough and considered review.  The comments were very helpful and allowed us to improve the paper.

In addition to the line items suggested by the referee, we have made the following substantive changes.

\begin{enumerate}
\item We have added a reference to Fintzen's paper as suggested, and provided a more precise condition on $p$ (1,2).
\item We added precise references to the work of Lusztig that we depend on (11,20,22).
\item We replaced $p$-adic field with local field as suggested (15).
\item We agree that it would be nice if the genericity condition on $\phi_i$ could be translated to an illuminating condition on $\varphi_i$.  Unfortunately, while the condition can certainly be translated to the Greenberg transform side of the story, it doesn't seem to provide any insight.  We have therefore not addressed the referee's suggestion (16).
\item We have added a reference to a paper of Yu (18).
\item In addressing (19,20), we realized that hypothesis \textbf{H0} was unnecessary (it resulted from our misreading of one of Lusztig's results) so we have removed it.  We have added two paragraphs illustrating hypothesis \textbf{H1(b)}, giving examples both of representations that do not satisfy it and a broad class of representations that do discussed in a paper of Kaletha.
\item We added a reference to Hakim's paper in the introduction (24).
\end{enumerate}

\end{document}

### 2020-10-20 Geometric types.tex
Remove todo notes, add referee response

\begin{abstract}
We show that some types for supercuspidal representations of tamely ramified $p$-adic groups that appear in Jiu-Kang Yu's work are geometrizable.
To do so, we define a function-sheaf dictionary for one-dimen\-sion\-al characters of arbitrary smooth group schemes over finite fields.  
In previous work we considered the case of commutative smooth group schemes and found that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  
In this paper we give a modification of the category of character sheaves that remedies this defect, and is also extensible to non-commutative groups.  
We then use these \emph{commutative character sheaves} to geometrize the linear characters that appear in the types introduced by Jiu-Kang Yu, assuming that the character vanishes on a certain derived subgroup.
To define \emph{geometric types}, we combine commutative character sheaves with Gurevich and Hadani's geometrization of the Weil representation and Lusztig's character sheaves.
\end{abstract}
\todo{Since Yu's article ``Construction of tame supercuspidal representation'' [20] is a key ingredient, I would like to inform the authors that, as recently noticed by Loren Spice, Proposition 14.1 and Theorem 14.2 in [20] are not always true (they rely on a misprinted statement in a reference). An alternative proof of Yu's main results which avoids these steps was provided by Jessica Fintzen in ``On the construction of tame supercuspidal representations'', arXiv:1908.09819. I think that it should be mentioned, even if does not affect the results of manuscript.}

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.3}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho_{i+1}$ & $\,^\circ \rho_{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3} \\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%
\todo{the quasicharacter $\varphi^i$ is defined to be the restriction to $^\circ K_i$ of the quasicharacter $\phi_i$ which appears in [20].  Yu assumes that $\phi^i$ is $G^{i+1}$ generic.  It would be interesting to see if $G^{i+1}$-genericity property could be translated into a condition of $\varphi^i$, viewed as a quasicharacter of $\underline{G}^i(R)$ as in M2.}

@@

We may now revisit the ingredients in the construction of the representation $\rho$ of $\underline{G}(R)$ along the lines indicated by Yu and recalled in Section~\ref{ssec:review}.
\begin{enumerate}
\labitem{M0}{M0}
The compact groups $\oK^i$ appearing in \ref{Y0} have been replaced by the smooth group schemes $\underline{G}^i$.
\labitem{M1}{M1}
The continuous representation $\orho^0$ of $\oK^0$ appearing in \ref{Y1} may be interpreted as a representation of $\underline{G}^0(R)$ obtained by inflation along $\underline{G}^0(R) \to \underline{G}^0(\Fq)$ from a representation $\varrho_0$ of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\varrho_0$ is itself obtained by pulling back a representation $\varrho_0^\reductive$; this is a consequence of \cite{Yu:models}*{\S 10.5 (i)}.
%\todo{why can $\varrho_0$ be obtained by pulling back a representation $\varrho_0^\reductive$} along the $\Fq$-points of the quotient $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$ 
\labitem{M2}{M2} The quasicharacters $\varphi^i$ appearing in \ref{Y2} are now quasicharacters of $\underline{G}^i(R)$, for $i=0, \ldots, d$.
In fact, if $r_i$ is the level of $\varphi^i$, as it appears in Section~\ref{ssec:quasicharacters}, then $\varphi^i$ is obtained by pulling back a quasicharacter $\varphi_i\ceq \varphi^i_{r_i}$ of the smooth group scheme $G_i\ceq G^i_{r_i} \ceq \Gr_{r_i}^{R}(\underline{G}^i)$ along $G^i \to G_i$.
\labitem{M3}{M3}
Diagram \eqref{eq:semiprod} in \ref{Y3} is now replaced by the following diagram of smooth group schemes over $R$.
\begin{equation}\label{eq:pimodel}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && \underline{G}^{i}\times_{\underline{G}} \underline{J}^{i+1}\arrow{d} && \\
1 \arrow{r} & \underline{J}^{i+1} \arrow{r} & \underline{G}^i \ltimes \underline{J}^{i+1} \arrow{d} \arrow{r} & \underline{G}^i \arrow{r} & 1\\
&& \underline{G}^{i+1} \arrow{d} && \\
&& 1 &&
\end{tikzcd}
\end{equation}
\labitem{M4}{M4}
The representation $h_{i+1}^*(W_{i+1})$ appearing in \ref{Y4} is now obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(R) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}

@@

As we recalled in Section~\ref{ssec:review}, a Yu type datum consists of compact groups $\oK^i$, a representation $\orho^0$ of $\oK^0$ and quasicharacters $\varphi^i$ of $\oK^i$, for $i=0,\ldots, d$; see \ref{Y0}, \ref{Y1} and \ref{Y2}.
In Section~\ref{ssec:review} we also saw how a Yu type datum determines smooth group schemes $\underline{G}^i$, a representation $\varrho_0\red$ of the reductive quotient $(\underline{G}^0)\red_{k}$ of the special fibre $\underline{G}^0_{k}$ of $\underline{G}^0$, and quasicharacters $\varphi_i$ of the group of $k$-rational points on the level-$r_i$ Greenberg transform $G_i$ of $\underline{G}^i$; see \ref{M0}, \ref{M1} and \ref{M2}.
%
Theorem~\ref{thm:geotypes} places the following conditions on the Yu type datum:
\begin{enumerate}
%\labitem{H0}{H0} the geometric component group of $(\underline{G}^0)\red_{k}$ is cyclic;
\labitem{H1}{H1} for each $i=0,\ldots, d$, either
\begin{enumerate}
\labitem{H1(a)}{H1a} $G_i$ is reductive, or % with cyclic geometric component group, or
\labitem{H1(b)}{H1b} the quasicharacter $\varphi_i$ of $G_i$ is trivial on  $G_{i,\operatorname{der}}(k)$. \todo{It would be worth to provide an example of supercuspidal representation (with positive depth) of a non-commutative p-adic group for which the hypothesis H1(b) is satsified}
\end{enumerate}
\end{enumerate}
%Hypothesis~\ref{H0} will allow us to use Lusztig's character sheaves to geometrize the character of the representation of $\orho^0$.
Hypothesis~\ref{H1a} may be used to geometrize each quasicharacter $\varphi_i$ using $\mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G_i}(G_i;\EE)$.
Alternatively, Hypothesis~\ref{H1b} allows us to use Proposition~\ref{prop:quasicharacters} to geometrize each quasicharacter $\varphi_i$ using local systems on $G_i$, or more precisely, using commutative character sheaves on $G_i$.
Using Sections~\ref{ssec:equivariant} and \ref{ssec:LCS} we see that, in both cases, the resulting geometrization may be interpreted as an element in $\mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G_i}(G_i;\EE)$.
Note that when Hypothesis~\ref{H1b} applies, it provides a considerably simpler geometrization than when Hypothesis~\ref{H1a} applies.

We give examples illustrating Hypothesis \ref{H1b}: first,
an example of a quasicharacter that does not vanish on $G_{i, \der}(k)$,
and then a class of Yu data where it is always satisfied.
Let $G = \SL_2$ over $\QQ_2$.
Then the hyperspecial maximal compact $\oK^0$ reduces to $\SL_2(\ZZ/2^{d+1}\ZZ)$ at depth $d$.
The derived subgroup of $\SL_2(\ZZ/2\ZZ)$ has index $2$, while it has index $4$ in positive depth.
There is therefore a character $\varphi^0$ of $K$ of order $4$ and depth $1$
factoring through the quotient of $\SL_2(\ZZ/4\ZZ)$ by its derived subgroup.
Taking the trivial representation for $\orho^0$ gives a Yu datum not satisfying \ref{H1b}. \todo{Does Yu require tameness?  Can we say that this is an example if so?}

@@

\begin{proof}
%For every $i=0, \ldots ,d$, set $G^i = \Gr_{R}(\underline{G}^i)$.
Recall that $G^i(\Fq) = \underline{G}^i(R) = \oK^i$, canonically.
Let $r_i$ be the level of $\varphi^i$ as defined in Section~\ref{ssec:quasicharacters}. 
%\todo{the precise reference in the series of Lusztig articles on character sheaves on disconnected groups must be provided.}
Set $r = \max\{ r_i \tq i=0, \ldots ,d\}$.

@@

We now define $\cs{A}^i_r  \in \mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G^i_r}$ for $i=0,\ldots ,d$ recursively, following the construction of the representations $\orho^i$, as reviewed in Section~\ref{ssec:review}.
First, set $\cs{A}_r^0 = A_r^0$ and note that \eqref{eqn:tracerho} commutes with $\trace(\orho^0)_r$ replaced by $\trFrob{\cs{A}_r^0}$.
Using Hypothesis~\ref{H1}, let $\cs{L}_i$ be the geometrization of the quasicharacter $\varphi_i \ceq \varphi^i_{r_i}$ appearing in \ref{M2}.
If \ref{H1a} applies, then $\cs{L}_i \in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_i}(G_i;\EE)$, using \cite{lusztig:disconnected4}*{Thm. 21.21};
if \ref{H1b} applies, then $\cs{L}_i\in \CCS(G_i)$, using Proposition~\ref{prop:quasicharacters}. 
%\todo{provide exact reference}
In either case, $\cs{L}_i\in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_i}(G_i;\EE)$ and
\[
\trFrob{\cs{L}_i} = \varphi_i.
\]
For each $i$, let $\cs{L}^i_{r}$ be the pull-back of $\cs{L}_i$ along $G^i_{r} \to G_i$; then $\cs{L}^i_{r}$ is a linear character sheaf and
\[
\trFrob{\cs{L}^i_{r}} = \varphi^i_{r}.
\]
Now, suppose $\cs{A}^i_r$ on $G_r^i$ is defined such that
\[
\begin{tikzcd}
G^i(\Fq) \arrow{rr}{\trace(\orho^i)} \arrow{rd} && \EE\\
& G_r^i(\Fq) \arrow{ru}[swap]{\trFrob{\cs{A}_r^i}} & 
\end{tikzcd}
\]
commutes.
Applying the Greenberg functor $\Gr^{R}_r$ to \eqref{eq:pimodel} gives
\begin{equation}\label{eq:pi}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && G_r^{i}\times_{G_r} J_r^{i+1} \arrow{d} && \\
1 \arrow{r} & J_r^{i+1} \arrow{r} & G_r^i \ltimes J_r^{i+1} \arrow{d}{\pi_r^{i+1}} \arrow{r}{p_r^{i}} & G_r^i \arrow{r} & 1\\
&& G_r^{i+1} \arrow{d} && \\
&& 1 && 
\end{tikzcd}
\end{equation}
where $J_r^{i+1} \ceq \Gr^{R}_r(\underline{J}^{i+1})$ and $G_r^{i} \ceq \Gr^{R}_r(\underline{G}^{i})$.
By \cite{bertapelle-gonzales:Greenberg}*{Prop 14.2}, the sequences are exact.
Consider $\cs{B}_r^{i+1}\in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_r^{i}\ltimes J_r^{i+1}}(G_r^{i}\ltimes J_r^{i+1};\EE)$ defined by
\[
\cs{B}_r^{i+1} \ceq \cs{K}_r^{i+1} \otimes (p_r^{i})^*(\cs{A}_r^{i}\otimes \cs{L}_r^{i}).
\]
Comparing with \eqref{eq:tensor}, we see that $\trFrob{\cs{B}_r^{i+1}}$ is precisely the function obtained by factoring the character of $\orho^{i+1}$ through $(\underline{G}^{i}\ltimes \underline{J}^{i+1})(R) \to (\underline{G}^{i}\ltimes \underline{J}^{i+1})(R_r)$ using the canonical identification $(G_r^{i}\times_{G_r} J_r^{i+1})(\Fq) =  (\underline{G}^{i}\ltimes \underline{J}^{i+1})(R_r)$. 
In particular, $\trFrob{\cs{B}^{i+1}_r}$ is constant on $(G_r^{i}\times_{G_r} J_r^{i+1})(\Fq)$, taking the value $\dim \orho^{i+1}$.
With reference to the morphism $\pi_r^{i+1} : G_r^i \ltimes J_r^{i+1} \to G_r^{i+1}$ from \eqref{eq:pi}, define 
\[
\cs{C}_r^{i+1} \ceq (\pi_r^{i+1})_! (\cs{B}_r^{i+1}).
\]
Then $\cs{C}_r^{i+1}\in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_r^{i+1}}(G_r^{i+1};\EE)$ and
\begin{eqnarray*}
\trFrob{\cs{C}^{i+1}_r}(x)
&=& \sum_{y\in (\pi_r^{i+1})^{-1}(x)}  \trFrob{\cs{B}_r^{i+1}}(y).
\end{eqnarray*}
Since $\trFrob{\cs{B}^{i+1}_r}$ is constant on $(G_r^{i}\times_{G_r} J_r^{i+1})(\Fq)$, it follows that 
\[
\trFrob{\cs{C}^{i+1}_r} = n \trFrob{\cs{B}^{i+1}_r}
\]
on $G_r^{i+1}(\Fq)$ with $n$ equal to the product of $\# (G_r^{i}\times_{G_r} J_r^{i+1})(\Fq)$ and $\dim \orho^{i+1}$.
Let $\cs{A}_r^{i+1}$ be the element of $\mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G_r^{i+1}}(G_r^{i+1};\EE)$ given by $\cs{A}_r^{i+1} = \frac{1}{n} \cs{C}_r^{i+1}$. 
%
This completes the inductive definition of $\cs{A}_r^i$ so that the following diagram commutes.
\[
\begin{tikzcd}
G^{i+1}(\Fq) \arrow{rr}{\trace(\orho^{i+1})} \arrow{rd} && \EE\\
& G_r^{i+1}(\Fq)\arrow{ru}[swap]{{\trFrob{\cs{A}_r^{i+1}}} } & 
\end{tikzcd}
\]

@@

\todo{It would be worth to quote also the recent article by Jeffrey Hakim ``Constructing tame
supercuspidal representations'' (Represent. Theory 22, 2018), which revisits Yu's construction in an
interesting manner.}

>>>>

\begin{abstract}
We show that some types for supercuspidal representations of tamely ramified $p$-adic groups that appear in Jiu-Kang Yu's work are geometrizable.
To do so, we define a function-sheaf dictionary for one-dimen\-sion\-al characters of arbitrary smooth group schemes over finite fields.  
In previous work we considered the case of commutative smooth group schemes and found that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  
In this paper we give a modification of the category of character sheaves that remedies this defect, and is also extensible to non-commutative groups.  
We then use these \emph{commutative character sheaves} to geometrize the linear characters that appear in the types introduced by Jiu-Kang Yu, assuming that the character vanishes on a certain derived subgroup.
To define \emph{geometric types}, we combine commutative character sheaves with Gurevich and Hadani's geometrization of the Weil representation and Lusztig's character sheaves.
\end{abstract}

@@

\begin{table}[ht]
\caption{Notation conversion chart.}
\begin{spacing}{1.3}
\begin{tabular}{| c|l | l | }
\hline
\text{this paper} & \text{Jiu-Kang Yu, {\it Construction of tame}} & \cite{yu:01a} \\
 & \text{{\it supercuspidal representations}} &  \\
\hline
$\oK^0$ & $\,^\circ K^0 = G^0(F)_y$ & \cite{yu:01a}*{\S 15} \\
$\oK^{i+1}$ & $\,^\circ K^{i+1} = (\,^\circ K^0) \vec{G}^{(i+1)}(F)_{y,(0, s_0, \ldots, s_{i})}$ & \cite{yu:01a}*{\S 15} \\
$\orho^0$ & $\,^\circ \rho^0$ & \cite{yu:01a}*{\S 15} \\
$\orho_{i+1}$ & $\,^\circ \rho_{i+1}$ &  \cite{yu:01a}*{\S 15} \\
$\varphi^i$ & $\phi_i\vert_{\,^\circ K^i }$ & \cite{yu:01a}*{\S 3} \\
$J_{i+1}$ & $J^{i+1} = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}$ & \cite{yu:01a}*{\S 3} \\
$V_{i+1}$ & $J^{i+1}/J^{i+1}_+ = (G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ (G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}$ & \cite{yu:01a}*{\S 3} \\ 
$V_{i+1}^\sharp$ & $(G^i,G^{i+1})(F)_{y, (r_i, s_i)}/ \ker(\widehat{\phi}_i\vert_{(G^i,G^{i+1})(F)_{y, (r_i, s_i^+)}})$ &  \cite{yu:01a}*{\S 4} \\
$Z_{i+1}$ & $\ker(V_{i+1}^\sharp\to V_{i+1})$ & \cite{yu:01a}*{\S 11} \\
$(f_{j+1}, j_{i+1})$ & $(f,j)$ & \cite{yu:01a}*{\S 11} \\
$\langle\ ,\ \rangle_{i+1}$ & $\langle\ ,\ \rangle$ & \cite{yu:01a}*{\S 11} \\
\hline
\end{tabular}
\end{spacing}
\label{table:notation}
\end{table}%

@@

We may now revisit the ingredients in the construction of the representation $\rho$ of $\underline{G}(R)$ along the lines indicated by Yu and recalled in Section~\ref{ssec:review}.
\begin{enumerate}
\labitem{M0}{M0}
The compact groups $\oK^i$ appearing in \ref{Y0} have been replaced by the smooth group schemes $\underline{G}^i$.
\labitem{M1}{M1}
The continuous representation $\orho^0$ of $\oK^0$ appearing in \ref{Y1} may be interpreted as a representation of $\underline{G}^0(R)$ obtained by inflation along $\underline{G}^0(R) \to \underline{G}^0(\Fq)$ from a representation $\varrho_0$ of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\varrho_0$ is itself obtained by pulling back a representation $\varrho_0^\reductive$; this is a consequence of \cite{Yu:models}*{\S 10.5 (i)}.
\labitem{M2}{M2} The quasicharacters $\varphi^i$ appearing in \ref{Y2} are now quasicharacters of $\underline{G}^i(R)$, for $i=0, \ldots, d$.
In fact, if $r_i$ is the level of $\varphi^i$, as it appears in Section~\ref{ssec:quasicharacters}, then $\varphi^i$ is obtained by pulling back a quasicharacter $\varphi_i\ceq \varphi^i_{r_i}$ of the smooth group scheme $G_i\ceq G^i_{r_i} \ceq \Gr_{r_i}^{R}(\underline{G}^i)$ along $G^i \to G_i$.
\labitem{M3}{M3}
Diagram \eqref{eq:semiprod} in \ref{Y3} is now replaced by the following diagram of smooth group schemes over $R$.
\begin{equation}\label{eq:pimodel}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && \underline{G}^{i}\times_{\underline{G}} \underline{J}^{i+1}\arrow{d} && \\
1 \arrow{r} & \underline{J}^{i+1} \arrow{r} & \underline{G}^i \ltimes \underline{J}^{i+1} \arrow{d} \arrow{r} & \underline{G}^i \arrow{r} & 1\\
&& \underline{G}^{i+1} \arrow{d} && \\
&& 1 &&
\end{tikzcd}
\end{equation}
\labitem{M4}{M4}
The representation $h_{i+1}^*(W_{i+1})$ appearing in \ref{Y4} is now obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(R) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}

@@

As we recalled in Section~\ref{ssec:review}, a Yu type datum consists of compact groups $\oK^i$, a representation $\orho^0$ of $\oK^0$ and quasicharacters $\varphi^i$ of $\oK^i$, for $i=0,\ldots, d$; see \ref{Y0}, \ref{Y1} and \ref{Y2}.
In Section~\ref{ssec:review} we also saw how a Yu type datum determines smooth group schemes $\underline{G}^i$, a representation $\varrho_0\red$ of the reductive quotient $(\underline{G}^0)\red_{k}$ of the special fibre $\underline{G}^0_{k}$ of $\underline{G}^0$, and quasicharacters $\varphi_i$ of the group of $k$-rational points on the level-$r_i$ Greenberg transform $G_i$ of $\underline{G}^i$; see \ref{M0}, \ref{M1} and \ref{M2}.
%
Theorem~\ref{thm:geotypes} places the following conditions on the Yu type datum:
\begin{enumerate}
%\labitem{H0}{H0} the geometric component group of $(\underline{G}^0)\red_{k}$ is cyclic;
\labitem{H1}{H1} for each $i=0,\ldots, d$, either
\begin{enumerate}
\labitem{H1(a)}{H1a} $G_i$ is reductive, or % with cyclic geometric component group, or
\labitem{H1(b)}{H1b} the quasicharacter $\varphi_i$ of $G_i$ is trivial on  $G_{i,\operatorname{der}}(k)$.
\end{enumerate}
\end{enumerate}
%Hypothesis~\ref{H0} will allow us to use Lusztig's character sheaves to geometrize the character of the representation of $\orho^0$.
Hypothesis~\ref{H1a} may be used to geometrize each quasicharacter $\varphi_i$ using $\mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G_i}(G_i;\EE)$.
Alternatively, Hypothesis~\ref{H1b} allows us to use Proposition~\ref{prop:quasicharacters} to geometrize each quasicharacter $\varphi_i$ using local systems on $G_i$, or more precisely, using commutative character sheaves on $G_i$.
Using Sections~\ref{ssec:equivariant} and \ref{ssec:LCS} we see that, in both cases, the resulting geometrization may be interpreted as an element in $\mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G_i}(G_i;\EE)$.
Note that when Hypothesis~\ref{H1b} applies, it provides a considerably simpler geometrization than when Hypothesis~\ref{H1a} applies.

We give examples illustrating Hypothesis \ref{H1b}: first,
an example of a quasicharacter that does not vanish on $G_{i, \der}(k)$,
and then a class of Yu data where it is always satisfied.
Let $G = \SL_2$ over $\QQ_2$.
Then the hyperspecial maximal compact $\oK^0$ reduces to $\SL_2(\ZZ/2^{d+1}\ZZ)$ at depth $d$.
The derived subgroup of $\SL_2(\ZZ/2\ZZ)$ has index $2$, while it has index $4$ in positive depth.
There is therefore a character $\varphi^0$ of $K$ of order $4$ and depth $1$
factoring through the quotient of $\SL_2(\ZZ/4\ZZ)$ by its derived subgroup.
Taking the trivial representation for $\orho^0$ gives a datum not satisfying \ref{H1b}.

@@

\begin{proof}
%For every $i=0, \ldots ,d$, set $G^i = \Gr_{R}(\underline{G}^i)$.
Recall that $G^i(\Fq) = \underline{G}^i(R) = \oK^i$, canonically.
Let $r_i$ be the level of $\varphi^i$ as defined in Section~\ref{ssec:quasicharacters}. 
Set $r = \max\{ r_i \tq i=0, \ldots ,d\}$.

@@

We now define $\cs{A}^i_r  \in \mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G^i_r}$ for $i=0,\ldots ,d$ recursively, following the construction of the representations $\orho^i$, as reviewed in Section~\ref{ssec:review}.
First, set $\cs{A}_r^0 = A_r^0$ and note that \eqref{eqn:tracerho} commutes with $\trace(\orho^0)_r$ replaced by $\trFrob{\cs{A}_r^0}$.
Using Hypothesis~\ref{H1}, let $\cs{L}_i$ be the geometrization of the quasicharacter $\varphi_i \ceq \varphi^i_{r_i}$ appearing in \ref{M2}.
If \ref{H1a} applies, then $\cs{L}_i \in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_i}(G_i;\EE)$, using \cite{lusztig:disconnected4}*{Thm. 21.21};
if \ref{H1b} applies, then $\cs{L}_i\in \CCS(G_i)$, using Proposition~\ref{prop:quasicharacters}. 
In either case, $\cs{L}_i\in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_i}(G_i;\EE)$ and
\[
\trFrob{\cs{L}_i} = \varphi_i.
\]
For each $i$, let $\cs{L}^i_{r}$ be the pull-back of $\cs{L}_i$ along $G^i_{r} \to G_i$; then $\cs{L}^i_{r}$ is a linear character sheaf and
\[
\trFrob{\cs{L}^i_{r}} = \varphi^i_{r}.
\]
Now, suppose $\cs{A}^i_r$ on $G_r^i$ is defined such that
\[
\begin{tikzcd}
G^i(\Fq) \arrow{rr}{\trace(\orho^i)} \arrow{rd} && \EE\\
& G_r^i(\Fq) \arrow{ru}[swap]{\trFrob{\cs{A}_r^i}} & 
\end{tikzcd}
\]
commutes.
Applying the Greenberg functor $\Gr^{R}_r$ to \eqref{eq:pimodel} gives
\begin{equation}\label{eq:pi}
\begin{tikzcd}
\ && 1 \arrow{d} && \\
\ && G_r^{i}\times_{G_r} J_r^{i+1} \arrow{d} && \\
1 \arrow{r} & J_r^{i+1} \arrow{r} & G_r^i \ltimes J_r^{i+1} \arrow{d}{\pi_r^{i+1}} \arrow{r}{p_r^{i}} & G_r^i \arrow{r} & 1\\
&& G_r^{i+1} \arrow{d} && \\
&& 1 && 
\end{tikzcd}
\end{equation}
where $J_r^{i+1} \ceq \Gr^{R}_r(\underline{J}^{i+1})$ and $G_r^{i} \ceq \Gr^{R}_r(\underline{G}^{i})$.
By \cite{bertapelle-gonzales:Greenberg}*{Prop 14.2}, the sequences are exact.
Consider $\cs{B}_r^{i+1}\in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_r^{i}\ltimes J_r^{i+1}}(G_r^{i}\ltimes J_r^{i+1};\EE)$ defined by
\[
\cs{B}_r^{i+1} \ceq \cs{K}_r^{i+1} \otimes (p_r^{i})^*(\cs{A}_r^{i}\otimes \cs{L}_r^{i}).
\]
Comparing with \eqref{eq:tensor}, we see that $\trFrob{\cs{B}_r^{i+1}}$ is precisely the function obtained by factoring the character of $\orho^{i+1}$ through $(\underline{G}^{i}\ltimes \underline{J}^{i+1})(R) \to (\underline{G}^{i}\ltimes \underline{J}^{i+1})(R_r)$ using the canonical identification $(G_r^{i}\times_{G_r} J_r^{i+1})(\Fq) =  (\underline{G}^{i}\ltimes \underline{J}^{i+1})(R_r)$. 
In particular, $\trFrob{\cs{B}^{i+1}_r}$ is constant on $(G_r^{i}\times_{G_r} J_r^{i+1})(\Fq)$, taking the value $\dim \orho^{i+1}$.
With reference to the morphism $\pi_r^{i+1} : G_r^i \ltimes J_r^{i+1} \to G_r^{i+1}$ from \eqref{eq:pi}, define 
\[
\cs{C}_r^{i+1} \ceq (\pi_r^{i+1})_! (\cs{B}_r^{i+1}).
\]
Then $\cs{C}_r^{i+1}\in \mathsf{K}_{\EE}\mathsf{D}^b_{c,G_r^{i+1}}(G_r^{i+1};\EE)$ and
\begin{eqnarray*}
\trFrob{\cs{C}^{i+1}_r}(x)
&=& \sum_{y\in (\pi_r^{i+1})^{-1}(x)}  \trFrob{\cs{B}_r^{i+1}}(y).
\end{eqnarray*}
Since $\trFrob{\cs{B}^{i+1}_r}$ is constant on $(G_r^{i}\times_{G_r} J_r^{i+1})(\Fq)$, it follows that 
\[
\trFrob{\cs{C}^{i+1}_r} = n \trFrob{\cs{B}^{i+1}_r}
\]
on $G_r^{i+1}(\Fq)$ with $n$ equal to the product of $\# (G_r^{i}\times_{G_r} J_r^{i+1})(\Fq)$ and $\dim \orho^{i+1}$.
Let $\cs{A}_r^{i+1}$ be the element of $\mathsf{K}_{\bar\QQ}\mathsf{D}^b_{c,G_r^{i+1}}(G_r^{i+1};\EE)$ given by $\cs{A}_r^{i+1} = \frac{1}{n} \cs{C}_r^{i+1}$. 
%
This completes the inductive definition of $\cs{A}_r^i$ so that the following diagram commutes.
\[
\begin{tikzcd}
G^{i+1}(\Fq) \arrow{rr}{\trace(\orho^{i+1})} \arrow{rd} && \EE\\
& G_r^{i+1}(\Fq)\arrow{ru}[swap]{{\trFrob{\cs{A}_r^{i+1}}} } & 
\end{tikzcd}
\]

@@
